---
title: "Wrangling Data Exercise"
format: html
---

##About the data
Data is from Spiny Lobster collection along a reef off the the coast of South-Central California between 2012 and 2022. Authors include LTER PIs of the Santa Barbara Coastal LTER: Reed, Daniel C, and Miller, Robert J. The data is available here: [https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sbc.77.8]

##Setup

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

##Read and explore data
```{r}
lobster_abundance <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=f32823fba432f58f66c06b589b7efac6")

lobster_traps <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=66dd61c75bda17c23a3bce458c56ed84")
```

##Exercise 2
```{r}
unique(lobster_abundance$SIZE_MM)

lobster_abundance <- lobster_abundance %>%
  mutate(SIZE_MM = ifelse(SIZE_MM == -99999, NA, SIZE_MM))
```
```{r}
unique(lobster_abundance$SIZE_MM)
```

```
subset_data <- subset(lobster_abundance, SIZE_MM > 50)

```{r}
subset_data <- lobster_abundance %>%
  filter(SITE == "AQUE" & SIZE_MM > 70) %>%
  select(SITE, SIZE_MM)
```

##Exercise 3

```{r}
subset_traps <- lobster_traps %>%
  filter(is.na(abundance)) %>%
  distinct(sites)
```

##Exercise 6