---
title: "data-wrangling"
format: html
---

##Load Libraries
```{r}
library(dplyr)
library(tidyr)
library(readr)
```


##Load Data

bit.ly/data-wrangle-lesson-data
```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```
```{r}
unique(catch_original$Region)
```

```{r}
colnames(catch_original)
```

```{r}
head(catch_original)
```
```{r}
summary(catch_original)
```

summary(catch_original)
#Data Cleaning and Wrangling
#Pipe Operator

#Select outcome

Goal: remove the "All" and "notersRedCode' columns

```{r}
catch_data <- catch_original %>%
  select("Region", "Year", "Chinook", "Sockeye", "Coho", "Pink", "Chum")
```

Another way of doing the same thing 
```{r}
catch_original %>% select(-All, -notesRegCode)
```

```{r}
dplyr::glimpse(catch_data)
```

##Change column content

GOAL: Convert the Chinook catch values to numeric type and overwrite the old Chinook column

```{r}
catch_clean <-catch_data %>%
  mutate(Chinook = as.numeric(Chinook))
```
#Find which line is NA (and needs to be converted to #)
```{r}
which(is.as(catch_clean))
```

Rewrite this with the pipe operator
```{r}
catch_clean$Chinook %>% 
  is.na() %>% 
  which()
```
```{r}
catch_data[i,]
```

GOAL: change the value of the 'chinook' column to 1 if the value is equal to the letter I

```{r}
catch_clean<- catch_data %>%
  mutate(Chinook = 
           if_else(condition = Chinook == "I", 
                   true= "1",
                   false = Chinook),
         Chinook = as.numeric(Chinook))
head(catch_clean)
```


## pivot_longer and pivot_wider

GOAL: Transform data from wide format into a long format

```{r}
catch_long <-catch_clean %>%
  #pivot longer all columns except Region and Year
  pivot_longer(
    cols = -c(Region, Year, Chinook),
    names_to = "species",
    values_to = "catch")
  )
```
## Renaming columns

GAOL Rename the 'catch' column to be called 'catch_thousands'

## Summary Statistics
bit.ly/3ppGUsQ
